---
title: "Analysis of the Malaysia Ageing and Retirement Survey (MARS) - Wave 1 and 2"
author: "Calvin"
date: "2025-01-01"
output: html_document
---

# 0. Preliminaries

```{r, warning = F, message = F}
rm(list=ls())

#install.packages(c("tidyverse","skimr","fixest","ivreg","rdrobust","margins","mgcv","boot","binsreg","AER","fwildclusterboot"))

library(tidyverse)
library(skimr)
library(fixest)     # For IV regressions
library(ivreg)      # Alternative IV regression package
library(rdrobust)   # Regression discontinuity
library(margins)    # Marginal effects
library(mgcv)       # Spline regressions
library(boot)
library(binsreg)
library(AER)

```

Import sav files

```{r}
wave1 <- haven::read_sav("C:/Users/Calvin/Documents/MARS Wave 1_new.sav")
wave2 <- haven::read_sav("C:/Users/Calvin/Documents/MARS Wave 2_new.sav")
#wave1 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 1_new.sav")
#wave2 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 2_new.sav")
```

# 1. Pre-analysis and wave merge

Check overlaps of IDs in Wave 1 and 2. Wave 1 uses "UNIQUEID", Wave 2 uses a whole different id coding system "sampleid", but luckily includes a variable "MARS1" that tracks the Wave 1 UNIQUEID of each Wave 2 respondent.

```{r}
# Rename variables 
wave1 <- wave1 %>% rename(id_wave1 = UNIQUEID)
wave2 <- wave2 %>% rename(id_wave2 = sampleid, id_wave1_reference = MARS1)

# ID overlap analysis
# Match wave2's MARS1 with wave1's UNIQUEID
common_ids <- intersect(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave1 <- setdiff(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave2 <- setdiff(wave2$id_wave1_reference, wave1$id_wave1)

# Summary of overlap
id_summary <- data.frame(
  Total_Wave1 = nrow(wave1),
  Total_Wave2 = nrow(wave2),
  Common_IDs = length(common_ids),
  Unique_Wave1 = length(unique_wave1),
  Unique_Wave2 = length(unique_wave2)
)

print(id_summary)
```

Pick out variables to join: Constant: want preserve only wave 1

Changing: want capture both waves

```{r}




```

Merge datasets using left join, by the MARS1 ids

```{r}
# Merge the datasets using the reference IDs for diagnostics
merged_data <- wave2 %>%
  filter(id_wave1_reference %in% wave1$id_wave1) %>%
  left_join(wave1, by = c("id_wave1_reference" = "id_wave1"), suffix = c("_wave2", "_wave1"))

```

## Replication of the Wave 1 analysis in 2023

```{r}

# Variable creation

#rename, recode variables and orient (higher values indicate higher wellbeing)
wave1 <- wave1 %>%
  mutate(
    depressed  = dplyr::recode(as.numeric(C301C), "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1),
    stress     = dplyr::recode(as.numeric(C301E), "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1),
    troublecon = dplyr::recode(as.numeric(C301B), "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1),
    isolation  = dplyr::recode(as.numeric(C301L), "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1)
  ) %>%
  rename(
    A205             = A205_recode,
    happy            = C301D,
    satisfaction     = C301G,
    activitiesfamily = C306B,
    volunteer        = C306D,
    read             = C306H,
    hobbies          = C306N,
    homemaint        = C306R,
    socialouting     = C306Q,
    exercise         = C306M
  )
         
#retirement variables
wave1 <- wave1 %>%
  mutate(retired = as.numeric(D101_new == 5))
  # full overlap between retired year listed and not working for pay currently

#money received
wave1 <- wave1 %>% 
  mutate(
    money_rec_child = rowSums(select(., starts_with("B104B_1_A_")), na.rm = TRUE), 
    money_rec_parent = rowSums(select(., starts_with("B203B_1_A_")), na.rm = TRUE), 
    money_rec_sib = rowSums(select(., starts_with("B304B_1_A_")), na.rm = TRUE),
    nosupport = as.numeric(money_rec_child < 1 & money_rec_parent < 1 & money_rec_sib < 1),
    income = rowSums(select(., starts_with("D102a_")), na.rm = TRUE),
    nomoney = as.numeric(nosupport == 1 & income < 1)
  )

#age variables and standardising ages
wave1 <- wave1 %>% 
  mutate(
    age            = as.numeric(Age),
    retireage      = as.numeric(age - (2018 - D101D_2)),
    agesq          = age^2,
    age3           = age^3,
    age4           = age^4,
    gender2        = as.numeric(A101RSex == 1),
    past55         = as.numeric(age >= 55),
    past60         = as.numeric(age >= 60),
    past54         = as.numeric(age >= 54),
    past56         = as.numeric(age >= 56),
    stdage         = age - 55,
    stdagesq       = stdage^2,        # equivalent to (age-55)^2
    stdage3        = stdage^3,        # equivalent to (age-55)^3
    stdpast55std   = as.numeric(stdage >= 0),
    stdpast54std   = as.numeric(age >= 54),  # indicator for age>=54
    stdpast56std   = as.numeric(age >= 56),  # indicator for age>=56
    stdpast60std   = as.numeric(age >= 60),  # indicator for age>=60
    int55age       = past55 * age,
    int55agesq     = past55 * agesq,
    int55age3      = past55 * age3,
    stdint55age    = ifelse(stdage >= 0, stdage, 0),
    stdint55agesq  = ifelse(stdage >= 0, stdagesq, 0),
    stdint55age3   = ifelse(stdage >= 0, stdage3, 0),
    int56age       = past56 * age,
    int56agesq     = past56 * agesq,
    int56age3      = past56 * age3
  )

#education variables
wave1 <- wave1 %>% 
  mutate(educ = case_when(
      A205 == 2 ~ 1,
      A205 == 3 ~ 2,
      A205 == 4 ~ 6,
      A205 == 5 ~ 9,
      A205 == 6 ~ 11,
      A205 == 7 | A205 == 8 ~ 12,
      A205 == 9 ~ 15,
      A205 == 10 ~ 17,
      TRUE ~ 0))
    
#outcome variables
wave1 <- wave1 %>% 
  mutate(
    mentalwb = (stress + depressed+isolation +troublecon + happy + satisfaction) / 6,
    socialwb = (activitiesfamily + volunteer + read + hobbies + homemaint + socialouting + exercise) / 7,
    totalwb = (stress + depressed+isolation + troublecon + activitiesfamily + volunteer + read + hobbies + homemaint + socialouting + exercise) / 13)

  
#other assets/wealth indicators for heterogeneity analysis
wave1 <- wave1 %>% 
  mutate(
    illnesses = rowSums(select(., starts_with("C104A_1_S_")), na.rm = TRUE),
    pen = as.numeric(E102_1_S_1 == 1 | E102_1_S_2 == 1 | E102_1_S_3 == 1),
    wealth = rowSums(select(., starts_with("F105B_A_")), na.rm = TRUE) + rowSums(select(., starts_with("F104B_A_")), na.rm = TRUE),
    married = as.numeric(A204_1 == 2))


# Subset
wave1_subset <- wave1 %>%
  filter(age <= 85) %>%
  select(id_wave1,retired,money_rec_child,money_rec_sib,money_rec_parent, retireage, age, gender2, past55, stdage, stdagesq, educ, mentalwb, socialwb, totalwb, stress, depressed, isolation, troublecon, happy, satisfaction, activitiesfamily, volunteer, read, hobbies, homemaint, socialouting, exercise, illnesses, pen, wealth,married)


#calculate bins for scatterplot
wave1_bins <- wave1_subset %>% 
  group_by(age) %>% 
  summarise(mean_retired = mean(retired, na.rm = TRUE))

```

```{r}
# produce a binscatter plot:
# - points: the average retired value by age
# - smooth line: a quadratic (qfit) linear model fitted on the original data
# - vertical line at the cutoff (rd(55))
ggplot(wave1_subset, aes(x = age, y = retired)) +
  stat_summary(fun = mean, geom = "point", binwidth = 5) +
  geom_smooth(data = dplyr::filter(wave1_subset, age <= 55 & age <70),
              method = "lm", formula = y ~ poly(x, 2, raw = TRUE),
              se = FALSE, linetype = "solid", colour = "blue") +
  geom_smooth(data = dplyr::filter(wave1_subset, age >= 55 & age >50),
              method = "lm", formula = y ~ poly(x, 2, raw = TRUE),
              se = FALSE, linetype = "solid", colour = "red") +
  geom_vline(xintercept = 55, linetype = "dashed") +
  labs(x = "Age", y = "Retirement", 
       title = "first stage: retirement on age at discontinuity 55")


#Total wellbeing model
model_total <- ivreg(totalwb2 ~ stdage + stdint55age + stdint55agesq + educ + hhlgender + illnesses + a200b + retired |
                       stdage + stdint55age + stdint55agesq + educ + hhlgender + illnesses + a200b + past55,
                       data = wave1_subset)

summary(model_total)





```
