---
title: "Analysis of the Malaysia Ageing and Retirement Survey (MARS) - Wave 1 and 2"
author: "Calvin"
date: "2025-01-01"
output: html_document
---

# 0. Preliminaries

```{r, warning = F, message = F}
rm(list=ls())

#install.packages(c("tidyverse","skimr","fixest","ivreg","rdrobust","margins","mgcv","boot"))

library(tidyverse)
library(skimr)
library(fixest)     # For IV regressions
library(ivreg)      # Alternative IV regression package
library(rdrobust)   # Regression discontinuity
library(margins)    # Marginal effects
library(mgcv)       # Spline regressions
library(boot)

```

Import sav files

```{r}
wave1 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 1_new.sav")
wave2 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 2_new.sav")
```

# 1. Pre-analysis and wave merge

Check overlaps of IDs in Wave 1 and 2. Wave 1 uses "UNIQUEID", Wave 2 uses a whole different id coding system "sampleid", but luckily includes a variable "MARS1" that tracks the Wave 1 UNIQUEID of each Wave 2 respondent.

```{r}
# Rename variables 
wave1 <- wave1 %>% rename(id_wave1 = UNIQUEID)
wave2 <- wave2 %>% rename(id_wave2 = sampleid, id_wave1_reference = MARS1)

# ID overlap analysis
# Match wave2's MARS1 with wave1's UNIQUEID
common_ids <- intersect(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave1 <- setdiff(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave2 <- setdiff(wave2$id_wave1_reference, wave1$id_wave1)

# Summary of overlap
id_summary <- data.frame(
  Total_Wave1 = nrow(wave1),
  Total_Wave2 = nrow(wave2),
  Common_IDs = length(common_ids),
  Unique_Wave1 = length(unique_wave1),
  Unique_Wave2 = length(unique_wave2)
)

print(id_summary)
```

Pick out variables to join: Constant: want preserve only wave 1

Changing: want capture both waves

```{r}




```

Merge datasets using left join, by the MARS1 ids

```{r}
# Merge the datasets using the reference IDs for diagnostics
merged_data <- wave2 %>%
  filter(id_wave1_reference %in% wave1$id_wave1) %>%
  left_join(wave1, by = c("id_wave1_reference" = "id_wave1"), suffix = c("_wave2", "_wave1"))

```

## Replication of the Wave 1 analysis in 2023, done in Stata code

```{r}


# Create retirement variables
wave1_1 <- wave1 %>%
  rename(A205 = A205_recode) %>%
  mutate(
    retired = ifelse(D101_new == 5, 1, 0),
    nosalary = ifelse(E102_1_S_1 != 3 & E102_1_S_2 != 3 & E102_1_S_3 != 3, 1, 0),
    retired2 = ifelse(D101_new == 5 & nosalary == 1, 1, 0),
    retired3 = ifelse(D101_new == 5 & D132_1A != 6, 1, 0),
    retired_notforced = ifelse(D101_new == 5 & D133 != 2, 1, 0),
    retired_forced = case_when(D133 == 2 ~ 1, D133 == 1 | D133 == 3 ~ 0, TRUE ~ NA_real_),
    retired_volun = case_when(D133 == 1 ~ 1, D133 == 2 | D133 == 3 ~ 0, TRUE ~ NA_real_),
    retired_volun3 = ifelse(D133 == 1 | D133 == 2, 1, 0),
    retired_forced2 = ifelse(D133 == 3 | D133 == 2, 1, 0),
    money_rec_child = rowSums(select(., starts_with("B104B_1_A_")), na.rm = TRUE),
    money_rec_parent = rowSums(select(., starts_with("B203B_1_A_")), na.rm = TRUE),
    money_rec_sib = rowSums(select(., starts_with("B304B_1_A_")), na.rm = TRUE),
    retired_nosupport = ifelse(D101_new == 5 & money_rec_child < 1 & money_rec_parent < 1 & money_rec_sib < 1, 1, 0),
    retireage = Age - (2018 - D101D_2),
    agesq = Age^2,
    age3 = Age^3,
    age4 = Age^4,
    educ = case_when(
      A205 == 2 ~ 1,
      A205 == 3 ~ 2,
      A205 == 4 ~ 6,
      A205 == 5 ~ 9,
      A205 == 6 ~ 11,
      A205 == 7 | A205 == 8 ~ 12,
      A205 == 9 ~ 15,
      A205 == 10 ~ 17,
      TRUE ~ 0
    ),
    C301all = C301D + C301G + C301O + C301P + C301Q + C301R + C301S + C301T - 
              (C301A + C301B + C301C + C301E + C301H + C301I + C301J + C301K + C301L + C301N),
    stdC301 = scale(C301all),
    gender2 = ifelse(A101RSex == 1, 0, 1),
   depressed = recode(as.numeric(C301C), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    stress = recode(as.numeric(C301E), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    troublecon = recode(as.numeric(C301B), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    isolation = recode(as.numeric(C301L), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    mentalwb = (stress + depressed) / 2,
    socialwb = (C306B + C306D + C306H + C306N + C306R + C306Q + C306M) / 7,
    totalwb = (stress + depressed + socialwb) / 9,
    illnesses = rowSums(select(., starts_with("C104A_1_S_")), na.rm = TRUE),
    past55 = ifelse(Age >= 55, 1, 0),
    stdage = Age - 55,
    stdagesq = stdage^2,
    pen = ifelse(E102_1_S_1 == 1 | E102_1_S_2 == 1 | E102_1_S_3 == 1, 1, 0),
    wealth = rowSums(select(., starts_with("F105B_A_")), na.rm = TRUE) + rowSums(select(., starts_with("F104B_A_")), na.rm = TRUE),
    married = ifelse(A204_1 == 2, 1, 0)
  )

# Regression Discontinuity Design (RDD)
rdd_model <- rdrobust(wave1_1$totalwb, wave1_1$stdage, c = 0, fuzzy = wave1_1$retired)

# Instrumental Variable Regression (IV)
iv_model <- ivreg(totalwb ~ stdage + educ + gender2 + illnesses + A200B | past55, data = wave1_1)

# Bootstrap Standard Errors
boot_iv <- boot(data = wave1_1, statistic = function(data, indices) {
  d <- data[indices, ]
  coef(ivreg(totalwb ~ stdage + educ + gender2 + illnesses + A200B | past55, data = d))
}, R = 1000)

# Visualizing Retirement by Age
ggplot(wave1_1, aes(x = Age, y = retired)) +
  geom_smooth(method = "loess", se = TRUE) +
  geom_vline(xintercept = 55, linetype = "dashed") +
  labs(title = "Retirement by Age", x = "Age", y = "Retirement Probability")

# Save results
summary(iv_model)
summary(rdd_model)

# Exporting Descriptive Statistics
descriptive_stats <- wave1_1 %>%
  select(Age, gender2, educ, married, illnesses, C101, B101, A200B, savings, totalwb, socialwb, mentalwb) %>%
  summary()

write.csv(descriptive_stats, "descriptive_stats.csv")

# Kernel Density Plots
ggplot(wave1_1, aes(x = Age)) +
  geom_density(fill = "blue", alpha = 0.3) +
  geom_vline(xintercept = 55, linetype = "dashed") +
  labs(title = "Kernel Density of Age", x = "Age", y = "Density")









```
