---
title: "Analysis of the Malaysia Ageing and Retirement Survey (MARS) - Wave 1 and 2"
author: "Calvin"
date: "2025-01-01"
output: html_document
---

# 0. Preliminaries

```{r, warning = F, message = F}
rm(list=ls())

#install.packages(c("tidyverse","skimr","fixest","ivreg","rdrobust","margins","mgcv","boot","binsreg","AER","fwildclusterboot"))

library(tidyverse)
library(skimr)
library(fixest)     # For IV regressions
library(ivreg)      # Alternative IV regression package
library(rdrobust)   # Regression discontinuity
library(margins)    # Marginal effects
library(mgcv)       # Spline regressions
library(boot)
library(binsreg)
library(AER)
library(fwildclusterboot)

```

Import sav files

```{r}
wave1 <- haven::read_sav("C:/Users/Calvin/Documents/MARS Wave 1_new.sav")
wave2 <- haven::read_sav("C:/Users/Calvin/Documents/MARS Wave 2_new.sav")
#wave1 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 1_new.sav")
#wave2 <- haven::read_sav("C:/Users/calvi/retirementMARS/MARS Wave 2_new.sav")
```

# 1. Pre-analysis and wave merge

Check overlaps of IDs in Wave 1 and 2. Wave 1 uses "UNIQUEID", Wave 2 uses a whole different id coding system "sampleid", but luckily includes a variable "MARS1" that tracks the Wave 1 UNIQUEID of each Wave 2 respondent.

```{r}
# Rename variables 
wave1 <- wave1 %>% rename(id_wave1 = UNIQUEID)
wave2 <- wave2 %>% rename(id_wave2 = sampleid, id_wave1_reference = MARS1)

# ID overlap analysis
# Match wave2's MARS1 with wave1's UNIQUEID
common_ids <- intersect(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave1 <- setdiff(wave1$id_wave1, wave2$id_wave1_reference)
unique_wave2 <- setdiff(wave2$id_wave1_reference, wave1$id_wave1)

# Summary of overlap
id_summary <- data.frame(
  Total_Wave1 = nrow(wave1),
  Total_Wave2 = nrow(wave2),
  Common_IDs = length(common_ids),
  Unique_Wave1 = length(unique_wave1),
  Unique_Wave2 = length(unique_wave2)
)

print(id_summary)
```

Pick out variables to join: Constant: want preserve only wave 1

Changing: want capture both waves

```{r}




```

Merge datasets using left join, by the MARS1 ids

```{r}
# Merge the datasets using the reference IDs for diagnostics
merged_data <- wave2 %>%
  filter(id_wave1_reference %in% wave1$id_wave1) %>%
  left_join(wave1, by = c("id_wave1_reference" = "id_wave1"), suffix = c("_wave2", "_wave1"))

```

## Replication of the Wave 1 analysis in 2023, done in Stata code

```{r}


# Create retirement variables
wave1_1 <- wave1 %>%
  rename(A205 = A205_recode) %>%
  mutate(
    retired = as.numeric(D101_new == 5),
    nosalary = as.numeric(E102_1_S_1 != 3 & E102_1_S_2 != 3 & E102_1_S_3 != 3),
    retired2 = as.numeric(D101_new == 5 & nosalary == 1),
    retired3 = as.numeric(D101_new == 5 & D132_1A != 6),
    retired_notforced = as.numeric(D101_new == 5 & D133 != 2),
    retired_forced = case_when(D133 == 2 ~ 1, D133 == 1 | D133 == 3 ~ 0, TRUE ~ NA_real_),
    retired_volun = case_when(D133 == 1 ~ 1, D133 == 2 | D133 == 3 ~ 0, TRUE ~ NA_real_),
    retired_volun3 = as.numeric(D133 == 1 | D133 == 2),
    retired_forced2 = as.numeric(D133 == 3 | D133 == 2),
    money_rec_child = rowSums(select(., starts_with("B104B_1_A_")), na.rm = TRUE),
    money_rec_parent = rowSums(select(., starts_with("B203B_1_A_")), na.rm = TRUE),
    money_rec_sib = rowSums(select(., starts_with("B304B_1_A_")), na.rm = TRUE),
    retired_nosupport = as.numeric(D101_new == 5 & money_rec_child < 1 & money_rec_parent < 1 & money_rec_sib < 1),
    retireage = as.numeric(Age - (2018 - D101D_2)),
    Age = as.numeric(Age),
    agesq = Age^2,
    age3 = Age^3,
    age4 = Age^4,
    educ = case_when(
      A205 == 2 ~ 1,
      A205 == 3 ~ 2,
      A205 == 4 ~ 6,
      A205 == 5 ~ 9,
      A205 == 6 ~ 11,
      A205 == 7 | A205 == 8 ~ 12,
      A205 == 9 ~ 15,
      A205 == 10 ~ 17,
      TRUE ~ 0
    ),
    C301all = C301D + C301G + C301O + C301P + C301Q + C301R + C301S + C301T - 
              (C301A + C301B + C301C + C301E + C301H + C301I + C301J + C301K + C301L + C301N),
    stdC301 = scale(C301all),
    gender2 = as.numeric(A101RSex == 1),
    depressed = recode(as.numeric(C301C), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    stress = recode(as.numeric(C301E), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    troublecon = recode(as.numeric(C301B), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    isolation = recode(as.numeric(C301L), `1` = 5, `2` = 4, `3` = 3, `4` = 2, `5` = 1),
    mentalwb = (stress + depressed) / 2,
    socialwb = (C306B + C306D + C306H + C306N + C306R + C306Q + C306M) / 7,
    totalwb = (stress + depressed + socialwb) / 9,
    illnesses = rowSums(select(., starts_with("C104A_1_S_")), na.rm = TRUE),
    past55 = as.numeric(Age >= 55),
    stdage = Age - 55,
    stdagesq = stdage^2,
    pen = as.numeric(E102_1_S_1 == 1 | E102_1_S_2 == 1 | E102_1_S_3 == 1),
    wealth = rowSums(select(., starts_with("F105B_A_")), na.rm = TRUE) + rowSums(select(., starts_with("F104B_A_")), na.rm = TRUE),
    married = as.numeric(A204_1 == 2)
  )



## first stage

# Ensure `wave1_1` has the needed variables
wave1_1 <- wave1_1 %>% filter(Age <= 85)

wave1_1_bins <- wave1_1 %>% 
  group_by(Age) %>% 
  summarise(mean_retired = mean(retired, na.rm = TRUE))

# produce a binscatter plot:
# - points: the average retired value by age
# - smooth line: a quadratic (qfit) linear model fitted on the original data
# - vertical line at the cutoff (rd(55))
ggplot(wave1_1, aes(x = Age, y = retired)) +
  stat_summary(fun = mean, geom = "point") +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, linetype = "solid") +
  geom_vline(xintercept = 55, linetype = "dashed") +
  labs(x = "Age", y = "Retirement")
print(p)


#Total wellbeing model
model_total <- ivreg(totalwb ~ stdage + retired | stdage + past55, data = wave1_1)

summary(model_total)





```
